---
title: "Lab9"
subtitle: "Importance sampling"
author: "Piotr Ginalski"
date: "2025-04-14"
output:
  pdf_document: default
---

As a first example, we are going to simulate random walk with different probability distribution of step.

```{r, monte carlo, basic experiment}
func <- rnorm
start_point <- 10

up_barrier <- 20
down_barrier <- 0

Z <- start_point
while(Z >= down_barrier && Z <= up_barrier){
  Z <- Z - func(1)
}
```

Now we are going to pack it in function.

```{r, creating functions}
simulation <- function(start_point, up_barrier, down_barrier, func){
  Z <- start_point
  while(Z >= down_barrier && Z <= up_barrier){
    Z <- Z - func(1)
  }
    return(ifelse(Z < down_barrier, 1, 0))
}
```

Let's pack it in function for calculating probabilities for different distributions.

```{r, function for many simulation}
calculate_probability <-  function(start_point, up_barrier, down_barrier, func, n_sim){
  outputs <- numeric(n_sim)
for (i in 1:n_sim){
  outputs[i] <- simulation(10, 20, 0, func)
}

paste("Estimator:", mean(outputs), "+-", 2*sd(outputs) / sqrt(n_sim))
}
```

After definition of our function, let's calculate some examples.

```{r, u = 0, var = 1}
gen_norm <- function(x) rnorm(1, 0, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

```{r, u = -0.1, var = 1}
gen_norm <- function(x) rnorm(1, -0.1, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

```{r, u = -0.2, var = 1}
gen_norm <- function(x) rnorm(1, -0.2, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

```{r, u = -0.3, var = 1}
gen_norm <- function(x) rnorm(1, -0.3, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

```{r, u = -0.4, var = 1}
gen_norm <- function(x) rnorm(1, -0.4, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

```{r, u = -0.5, var = 1}
gen_norm <- function(x) rnorm(1, 0.5, 1)
calculate_probability(10, 20, 0, gen_norm, 10000)
```

Now we are going to focus on importance sampling. Lets' clear an environment.

```{r, clearing env}
rm(list = ls())
```

```{r, importance sampling}


weighted_simulation <- function(start_point, up_barrier, down_barrier, func, mean){
  Z <- start_point
  while(Z >= down_barrier && Z <= up_barrier){
    Z <- Z - func(1)
  }
    return(ifelse(Z < down_barrier, 1, 0) * exp(2*mean*Z) * exp(-2*mean*start_point))
}



calculate_probability_weighted <-  function(start_point, up_barrier, down_barrier, func, mean, n_sim){
  outputs <- numeric(n_sim)
  for (i in 1:n_sim){
    outputs[i] <- weighted_simulation(10, 20, 0, func, mean)
  }

paste("Estimator:", mean(outputs), "+-", 2*sd(outputs) / sqrt(n_sim))
}
```

Let's use our function!

```{r}
mean <- 0
sd <- 1

gen_norm <- function(x) rnorm(1, mean, sd)
calculate_probability_weighted(10, 20, 0, gen_norm, mean, 10000)
```


```{r}
mean <- 0.1
sd <- 1

gen_norm <- function(x) rnorm(1, mean, sd)
calculate_probability_weighted(10, 20, 0, gen_norm, mean, 10000)
```


```{r}
mean <- 0.2
sd <- 1

gen_norm <- function(x) rnorm(1, mean, sd)
calculate_probability_weighted(10, 20, 0, gen_norm, mean, 10000)
```

```{r}
mean <- 0.3
sd <- 1

gen_norm <- function(x) rnorm(1, mean, sd)
calculate_probability_weighted(10, 20, 0, gen_norm, mean, 10000)
```

```{r}
mean <- 0.5
sd <- 1

gen_norm <- function(x) rnorm(1, mean, sd)
calculate_probability_weighted(10, 20, 0, gen_norm, mean, 10000)
```