---
title: "Laboratory 1"
author: "Piotr Ginalski"
subtitle: Basics of generating random variables
output:
  pdf_document: default
  html_notebook: default
---

Firstly, let's generate uniform random variables.
```{r generating uniform random variables}
n <- 12

set.seed(10)
x <- runif(n)
x
```
We can easily approximate normal distribution by CLT.
```{r generowanie normalnego}
normal_variable <- (sum(x) - n/2) / (sqrt(n/12))
normal_variable
```
And we have a function which returns realization of normal distribution.
```{r function for normal random variable}
normal <- function(){
  n <- 12
  x <- runif(n)
  normal_variable <- (sum(x) - n/2) / (sqrt(n/12))
  return(normal_variable)
}
normal()
```
Generating $m$ realizations.
```{r using function}
m <- 200
normal_vector <- vector()
for (i in 1:m){
  normal_vector <- c(normal_vector, normal())
}
```
We plot the result.
```{r plotting}
{
hist(normal_vector, prob = TRUE)
rug(normal_vector)
curve(dnorm, from = -3, to = 3, add = TRUE)
}
```
Plotting the empirical cumulative distribution function.
```{r empirical distribution}
ecdf(normal_vector)
{
plot(ecdf(normal_vector))
curve(pnorm, from = -3, to = 3, add = TRUE)
}
ks.test(normal_vector, pnorm)
```
Proceeding to another task and defining the Polya urn scheme
```{r generating bernoullis}
bern <- function(alfa, beta){
  x <- runif(1)
  return(as.integer(x < alfa / (alfa + beta)))
}

experiment <- function(alfa, beta, n){
X <- vector()
for (i in 1:n){
  Xn <- bern(alfa + sum(X), beta + length(X) - sum(X))
  X <- c(X, Xn)
}

S <- cumsum(X)
return(S)
}
```
Performing the Polya experiment.
```{r repeating experiment}
alfa <- 1
beta <- 1
n <- 2000
data_for_plot <- data.frame(x = 1:n)

for(i in 1:20){
  results <- experiment(alfa, beta, n)
  name <- paste0("experiment", i)
  data_for_plot[name] <- results / (1:n)
}

# Creating plot
matplot(data_for_plot$x, data_for_plot[, -1], type = "l", lty = 1,
        xlab = "Index", ylab = "Result/Index", main = "Results of experiments")
```

```{r generating beta}
generate_betas <- function(alfa, beta, n, m){
 betas <- vector()

  for(i in 1:m){
    results <- experiment(alfa, beta, n)
    betas <- c(betas, results[n]/n)
  }
 return(betas)
}

my_betas <- generate_betas(1, 1, 200, 200)
ks.test(my_betas, "pbeta", shape1 = 1, shape2 = 1)

my_betas <- generate_betas(3, 2, 200, 200)
ks.test(my_betas, "pbeta", shape1 = 3, shape2 = 2)

my_betas <- generate_betas(11, 1, 200, 200)
ks.test(my_betas, "pbeta", shape1 = 11, shape2 = 1)

my_betas <- generate_betas(0.5, 0.5, 200, 200)
ks.test(my_betas, "pbeta", shape1 = 0.5, shape2 = 0.5)

my_betas <- generate_betas(5, 31, 200, 200)
ks.test(my_betas, "pbeta", shape1 = 5, shape2 = 31)
```
Proceeding to the next task. Clearing the environment.
```{r clearing env}
rm(list = ls())
```
Performing exercise 1.8 from the script.
We consider the arcsine law and define the random variable T_n – the proportion of time we were positive during the games.
The function random_walk takes an argument 'distribution' which specifies the distribution to sample X_i from.
```{r definition Sn and Tn}
random_walk <- function(distribution, n, ...){
Sn <- cumsum(distribution(n, ...))
Sn <- c(0, Sn)
Tn <- cumsum(as.integer(Sn > 0)) / (1:(n+1))

return(list(Sn = Sn, Tn = Tn))
}


two_pointed <- function(n, alfa, beta){
  x <- runif(n)
  return(as.integer(x < alfa / (alfa + beta))*2 - 1)
}

normal_walk <- random_walk(rnorm, 500, mean = 0, sd = 1)
unif_walk <- random_walk(runif, 500, min = -1, max = 1)
bern_walk <- random_walk(two_pointed, 500, alfa = 1, beta = 1)

plot(normal_walk$Sn, type = "l", ylab = "Trajectory") + abline(h = 0) 
plot(unif_walk$Sn, type = "l", ylab = "Trajectory") + abline(h = 0)
plot(bern_walk$Sn, type = "l", ylab = "Trajectory") + abline(h = 0)


generate_Tn <- function(distribution, n, m, ...){
result <- vector()
for (i in (1:m)){
  walk <- random_walk(distribution, n, ...)
  result <- c(result, walk$Tn[length(walk$Tn)])
}
return(result)
}

realization_Tn <- generate_Tn(rnorm, 500, 50, mean = 0, sd = 1)
ks.test(realization_Tn, "pbeta", shape1 = 1/2, shape2 = 1/2)
```

Proceeding to the next task. Clearing the environment.
```{r clearing env 2}
rm(list = ls())
```

Creating discrete variables with fixed probabilities.
```{r}
argmin <- vector()

for (i in 1:10000){
W = rexp(3, rate = c(1, 2, 5))
argmin <- c(argmin, which.min(W))
}

table(argmin) / length(argmin)

# Function for generating discrete variables
generate_discretes <- function(n, freq){
  argmin <- vector()
  for (i in 1:n){
    W = rexp(length(freq), rate = freq)
    argmin <- c(argmin, which.min(W))
  }
  return(argmin)
}

n <- 1000
vector1 <- c(1,2,3)
vector2 <- c(1, 1, 10, 2, 3)
vector3 <- c(8, 12)

discrete_realization <- generate_discretes(n, vector1)
table(discrete_realization) / n

discrete_realization <- generate_discretes(n, vector2)
table(discrete_realization) / n

discrete_realization <- generate_discretes(n, vector3)
table(discrete_realization) / n

```
And finally, let's clear environment.
```{r clearing env 3}
rm(list = ls())
```

